### 2021-11-14 draft.tex
Deploying to gh-pages from @ gottacatchenall/thesis_proposal@d9eb006258265857374832c5e6edb4e2cad11213 🚀

As a result \emph{ecological forecasting} (\textbf{dietze?}), or
modeling how ecological systems will change over time, has as an
imperative to mitigate the effect of these changes on Earth's
ecosystems, their functioning, and the services they provide to humans.

An oft applied definition of the origin of ecology is ``the application
of the scientific method to natural history.'' Since its origin ecology
has been a descriptive science. This is a natural biproduct of the
immense variability of Earth's biosphere.

emerged to explain particular phenomena at particular scales.

@@

This term implicitly creates an analogy between predicting how
ecosystems will change in the future and weather forecasting. Use of
computational methods in NWP. Much as one would not aim to forecast the
weather to Quebec by applying Navier-Stokes. NWP has worked because it
incorporates information about data and meteorological processes
collected at difference scales into models that.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

Historically the term ``theory,'' as applied in the physical sciences,
refers to mathematical models. Many (although not all) of these models
refers to a equation describing how the value of an observable state of
the system,
\(\begin{bmatrix}x_1 \\ x_2 \\ \vdots \end{bmatrix} = \vec{x}\), changes
as a function of time.

differential equations (in continuous time, difference equations for
discrete time)

Ecological processes vary across more variables than the tools of
analytic models are suited for. As the number of variables in an
analytic model increases, so does the ability of the scientist to decern
clear relationships between them. Chaotic dynamics emerge from simple
analytic models, and .

Whether ecosystems actually exhibit chaotic behavior is a different
question.

Until the 20th century, no theory of the gravitational dynamics of more
than 2 bodies. Understanding the gravitational dynamics of more than two
planets with any reliability proved difficult. Using the same models
(diffeqs), how could we adaquetly predict ecosystems?

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

@@

\hypertarget{ch1-forecasting-the-phenological-uncoupling-of-a-plant-pollinator-network}{%
\section{CH1 Forecasting the phenological uncoupling of a
plant-pollinator
network}\label{ch1-forecasting-the-phenological-uncoupling-of-a-plant-pollinator-network}}

@@

@@

@@

\begin{itemize}
\tightlist
\item
  simulate species distribution and efficacy of detection given a set of
  observation points where the dist from observation site decays.
\item
  optimize set of repeated sampling locations L for a \emph{known}
  distribution D.
\item
  address SDM not being the territory
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  the missing link paper, turn this into optimizing with two different
  SDMs
\end{itemize}

>>>>

As a result \emph{ecological forecasting}, or building models to
estimate how ecological systems will change over time, has as an
imperative to mitigate the effect of these changes on Earth's
ecosystems, their functioning, and the services they provide to humans
(\textbf{dietze?}).

An oft applied definition of the origin of is ``the application of the
scientific method to natural history.'' Since its origin ecology has
been a descriptive science. This is a natural by-product of the immense
variability of Earth's biosphere. emerged to explain particular
phenomena at particular scales.

@@

Historically the term ``theory,'' as applied in the physical sciences,
refers to mathematical models, typically an equation describing how the
value of an observable state of the system, represented by a vector of
numbers \([x_1, x_2, \dots, x_n]^T = \vec{x}\) changes as over time.

\(\vec{x}(t)\) but instead to define how the state of \(\vec{x}\)
changes from one time to the next.

Because of its early success in the physical science, the led to
framework for bridging theory and data.

A large set of problems in ecology when aiming to confront
high-dimensional analytic models with data:

\begin{itemize}
\tightlist
\item
  Often assume long-run equilibrium.
\end{itemize}

Ecological processes vary across more variables than the tools of
analytic models are suited for.

As the number of variables in an analytic model increases, so does the
ability of the scientist to decern clear relationships between them, and
so does overfitting potential.

Curse of dimensionality--- Until the 20th century, no theory of the
gravitational dynamics of more than 2 bodies. Understanding the
gravitational dynamics of more than two planets with any reliability
proved difficult. Using the same models (diffeqs), how could we
adequately predict ecosystems?

Chaotic dynamics emerge from simple analytic models, and . Whether
ecosystems actually exhibit chaotic behavior is a different question.

The term \emph{ecological forecasting} implicitly creates an analogy
between predicting how ecosystems will change in the future and weather
forecasting. Use of computational methods in NWP. Much as one would not
aim to forecast the weather in Quebec by applying Navier-Stokes. NWP has
worked because it incorporates information about data and meteorological
processes collected at difference scales into models that.

Transition to simulation as the solution: shift toward approach of
building models that \emph{generate} data.

(resolving the semantic ambuity of what differentiates ``mechanistic''
vs ``phenomological'' models is out of scope for now). ----

@@

\hypertarget{ch1-forecasting-the-spatial-and-phenological-uncoupling-of-a-plant-pollinator-network}{%
\section{CH1 Forecasting the spatial and phenological uncoupling of a
plant-pollinator
network}\label{ch1-forecasting-the-spatial-and-phenological-uncoupling-of-a-plant-pollinator-network}}

@@

Two ways in which this network of interactions can become uncoupled:
spatial and temporal. Overlap in ranges and shifts in ranges.
Elevational gradient as proxy for range shifts

@@

System description: lots of data on \emph{Bombus} (bumblebees) and
wildflowers. Three different sites, (7/7/3) years each, each covering an
elevational gradient.

@@

Split the process into parts.

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  Building an interaction prediction model.
\item
  Make it spatial based on distributions.
\item
  Forecast distributions based on CMIP6.
\end{enumerate}

@@

\begin{itemize}
\item
  the missing link paper, turn this into optimizing with two different
  SDMs
\item
  relative abundance and its effect on false negative
\item
  non-independent associations in samples
\item
  simulate species distribution and efficacy of detection given a set of
  observation points where the dist from observation site decays.
\item
  optimize set of repeated sampling locations L for a \emph{known}
  distribution D.
\item
  address SDM not being the territory
\end{itemize}
